Kindly Bent to Free Us

### 2019-07-01 kindly/motivation.tex
Improve the region explanation.

A pool is parameterized by its content. Additional, the kind of the pool
depends on the content: Linear content implies
a linear pool, while unrestricted content implies an unrestricted pool.
The implementation uses a concurrent queue
as a bag of elements with atomic add and remove operations.
The \lstinline/Pool.create/ and \lstinline/Pool.consume/ allow to build or destroy
the pool by providing the necessary components for the elements
of the loop.
The main function of interest is the \lstinline/Pool.use/ function, which
allow to use an mutable borrow of the object. The function proceeds by removing
the element from the pool (or creating a fresh element),
passing it down to the \lstinline/f/ function, and then adding
it back to the pool.
We explicitly delimit the region on \cref{line:pool:region} to ensure that
the \lstinline/f/ function will not be able to capture \lstinline/&!o/.

\TODO{finish explanation}

\begin{figure}[!h]
  \centering
  \begin{subfigure}[t]{.45\linewidth}
\begin{lstlisting}
type ('a: 'k) pool : 'k = {
  spawn : unit -> 'a ;
  queue : 'a queue;
}

@@

let use { spawn ; queue } f =
  let o = match CQueue.pop &queue with
    | Some x -> x
    | None () -> spawn ()
  in
  let r = {| f &!o |} in(*@\label{line:pool:region}*)
  Queue.push o &queue;
  r
\end{lstlisting}
    \caption{Implementation}
    \label{impl:pool}
  \end{subfigure}~
  \begin{subfigure}[t]{.55\linewidth}
\begin{lstlisting}
type ('a: 'k) pool : 'k

create : (unit -> 'a) -> 'a pool
consume : ('a -> unit) -> 'a pool -> unit
use : ('a_1:'k_1),('a_2:'k_2),('k_1 < lin_r),('k_2 < lin_r) =>
  &('a_1 pool) -> (&!(aff_r,'a_1) -{'k}> 'a_2) -{'k_1}> 'a_2
\end{lstlisting}
    \caption{Interface}
    \label{intf:pool}

>>>>

A pool is parameterized by its content. The kind of the pool
depends on the content: linear content implies
a linear pool while unrestricted content implies an unrestricted pool.
The functions \lstinline/Pool.create/ and \lstinline/Pool.consume/ allow
to build or destroy
the pool by providing the corresponding functions for the elements
of the pool.
The main function is \lstinline/Pool.use/ which
allow to use an object from the pool.
\lstinline/Pool.use/ takes a shared borrow of a pool (in order to allow
for concurrent access) and a function. This function will receive
a mutable borrow of one arbitrary object of the pool. The typing ensure
that this borrow is not captured nor kept by the function.
This is done through a \emph{region} variable $r$ which annotates
the kinds. Here, the borrow is of type \lstinline/&!(aff_r,'a_1 )/, which indicates
a mutable borrow with an affine kind. The return type of \lstinline/f/ is bound
to be of kind \lstinline/'k_2 < aff_r/. Since this is a strict inequality,
types whose kind mention the region can not be returned.

The implementation presented in \cref{impl:pool} uses a concurrent queue
as a bag of elements with atomic add and remove operations.
The implementation for the \lstinline/Pool.create/ and \lstinline/Pool.consume/
functions are straightforward.
The function \lstinline/Pool.use/ first draws
an element from the pool (or creates a fresh element),
pass it down to the \lstinline/f/ function, and then add
it back to the pool.
We explicitly delimit the region on \cref{line:pool:region} to ensure that
the return value of \lstinline/f &!o/ do not capture \lstinline/&!o/.

\begin{figure}[!h]
  \centering
  \begin{subfigure}[t]{.45\linewidth}
\begin{lstlisting}
type ('a:'k) pool : 'k = {
  spawn : unit -> 'a ;
  queue : 'a CQueue.t;
}

@@

let use { spawn ; queue } f =
  let o = match CQueue.pop &queue with
    | Some x -> x
    | None () -> spawn ()
  in
  let r = {| f &!o |} in(*@\label{line:pool:region}*)
  Queue.push o &queue;
  r
\end{lstlisting}
    \caption{Implementation}
    \label{impl:pool}
  \end{subfigure}~
  \begin{subfigure}[t]{.55\linewidth}
\begin{lstlisting}
type ('a:'k) pool : 'k

create : (unit -> 'a) -> 'a pool
consume : ('a -> unit) -> 'a pool -> unit
use : ('a_1:'k_1),('a_2:'k_2),('k_1 < aff_r),('k_2 < aff_r) =>
  &('a_1 pool) -> (&!(aff_r,'a_1) -{lin}> 'a_2) -{'k_1}> 'a_2
\end{lstlisting}
    \caption{Interface}
    \label{intf:pool}
